Swarm and Evolutionary Based Algorithms used for Optimization

### 2023-01-19 latex/article.tex
header.tex done

\begin{abstract}
    In this paper, a study of evolution and swarm based algorithms is presented,
    using two classical engineering problems: \textit{Spring Tension} and \textit{Pressure Vessel
    Designs}. The test code for the problems was made using the \textit{Python Language},
    version 3.10 and uses \textit{MealPy} package, version 2.5.1,  to provide the algorithms.
    Thre algorithms were randomly chosen from a vaste list from MealPy algorithms:
    \textit{Evolutionary Programming (LevyEP)}, \textit{Evolution Strategies (OriginalES)} and
    \textit{Genetic Algorithm (BaseGA)} from \textit{evolutionary\_based} subpackage;
    \textit{Bees Algorithm (OriginalBeesA)}, \textit{Firefly Algorithm (OriginalFFA)} and
    \textit{Particle Swarm Optimization (OriginalPSO)} from \textit{swarm\_based} subpackage.
    Each problem was modeled using \textit{SciPy} package, with constraints implemented as
    \textit{penalty functions}. Each algorithm were optimized separately to extract the best
    solutions from each problem using the \textit{MealPy}'s \textit{Tuner} utility.
    The results, however, are dependant of algorithm and problem presented but from
    similarity test using Friedman's Chi-Squared test, it would be suffice to say that using any
    of the tested algorithm will produce similar results in these two optimization problems.
\end{abstract}

@@

\section{Metodology}
\label{sec:metodology}

\subsection{Optimimzation Problem Selection}
The two problems selected for this paper were \textit{Spring Tension Design} and \textit{Pressure Vessel Design}. Although it was simple
to choose the first two problems from the computational work statements, the choice was more than justified because these aroblems are well
known in the literature. Thus, the problem selection was driven by which has more that one source to compare results.

\subsection{Programming Language}

@@

\subsection{Algorithm Selection}

The algorithm selection was made in two steps: first, it was extracted simple version of the two classes (evolutionary and swarm based) and
then it was used a simple shuffle using

\subsection{Model Parameters}

For all models and problems, were selected the following parameters:

@@

\subsection{Model Tuning}

>>>>

\begin{abstract}
    In this paper, a study of evolution and swarm based algorithms is presented,
    using two classical engineering problems: \textit{Spring Tension} and \textit{Pressure Vessel
    Designs}. The test code for the problems was made using the \textit{Python Language},
    version 3.10 and uses \textit{MealPy} package, version 2.5.1,  to provide the algorithms.
    The algorithms were randomly chosen from a vast list from \textit{MealPy}'s algorithms:
    \textit{Evolutionary Programming (LevyEP)}, \textit{Evolution Strategies (OriginalES)} and
    \textit{Genetic Algorithm (BaseGA)} from \textit{evolutionary\_based} subpackage;
    \textit{Bees Algorithm (OriginalBeesA)}, \textit{Firefly Algorithm (OriginalFFA)} and
    \textit{Particle Swarm Optimization (OriginalPSO)} from \textit{swarm\_based} subpackage.
    Each problem was modeled using \textit{SciPy} package, with constraints implemented as
    \textit{penalty functions}. Each algorithm were optimized separately to extract the best
    solutions from each problem using the \textit{MealPy}'s \textit{Tuner} utility.
    The results, however, are dependant of algorithm and/or problem solved and the Friedman's chi
    squared test for similarity make it noticeable because, although the values for best fits are
    similar, running the same algorithm with different initial conditions does not converge to similar
    values.
\end{abstract}

@@

\section{Methodology}
\label{sec:methodology}

\subsection{Optimization Problem Selection}
\label{subsec:optimization-problem-selection}
The two problems selected for this paper were \textit{Spring Tension Design} and \textit{Pressure Vessel Design}. Although it was simple
to choose the first two problems from the computational work statements, the choice was more than justified because these aroblems are well
known in the literature.
Thus, the problem selection was driven by which has more than one source to compare results.

\subsubsection{Spring Tension Design}
\label{subsubsec:methodology-spring-tension-design}
From \textit{Nature-Inspired Metaheuristic Algorithms - Xin-She Yang} -
The design of a tension and compression spring is a well-known benchmark
optimization problem.
The main aim is to minimize the weight subject
to constraints on deflection, stress, surge frequency and geometry. It involves
three design variables: the wire diameter $x_1$ , coil diameter $x_2$ and
number/length of the coil $x_3$.
This problem can be summarized as:

\begin{equation}
    f (x) = x_1^2 x_2 (2 + x_3)
\end{equation}

subject to

\begin{equation}
    \begin{split}
        g_1(x) = \frac{x_2^3 x_3}{71785 x_1^4} \leq 0\\
        g_2(x) = \frac{4 x_2^2 - x_1 x_2}{12566 \left( x_1^3 x_2 - x_1^4\right)}\\
        + \frac{1}{5108 x_1^2} -1 \leq 0\\
        g_3(x) = 1 - \frac{140.45 x_1}{x_2^2 x_3} \leq 0\\
        g_4 = \frac{x_1 + x_2}{1.5} - 1 \leq 0
    \end{split}
\end{equation}

The original bounding limits of the variables, extracted from computational work
statements, are:

\begin{equation}
    \begin{split}
        0.05 \leq x_1 \leq 2.0\\
        0.25 \leq x_2 \leq 1.3\\
        2.0 \leq x_3 \leq 15.0
    \end{split}
\end{equation}

\subsubsection{Pressure Vessel Design}
\label{subsubsec:methodology-pressure-vessel-design}
From \textit{Solving Design of Pressure Vessel Engineering
Problem Using a Fruit Fly Optimization Algorithm - XIANTING KE et al} -
A pressure vessel design model involves four decision variables: $x_1$ is defined
thickness of the pressure vessel $T_s$ , $x_2$ stands for
thickness of the head $T_H$ , $x_3$ represents inner radius of the
vessel $R$ , and $x_4$ is on behalf of length of the vessel
barring head $L$ , the total variables described as $( x_1 , x_2 , x_3 , x_4 )$.
The objective function of the problem is to minimize the total cost, including
the cost of material, forming, and welding.
The general pressure vessel design optimization model is expressed as:

\begin{equation}
	\begin{split}
        f(x) = 0.6224 x_1 x_3 x_4 + 1.7781 x_2 x_3^2 \\
        + 3.1661 x_1^2 x_4 + 19.84 x_1^2 x_3
	\end{split}
\end{equation}

subject to:

\begin{equation}
    \begin{split}
        g_1(x) = - x_1 + 0.0193 x_3 \leq 0\\
        g_2(x) = - x_2 + 0.00954 x_3 \leq 0\\
        g_3(x) = - \pi x_3^2 x_4 -\frac{4}{3}\pi x_3^3 + 1296000 \leq 0\\
        g_4(x) = x_4 -240 \leq 0\\
    \end{split}
\end{equation}

The original bounding limits of the variables, extracted from computational work
statements, are:

\begin{equation}
    \begin{split}
        0 \leq x_1 \leq 99\\
        0 \leq x_2 \leq 99\\
        10 \leq x_3 \leq 200\\
        10 \leq x_4 \leq 200\\
    \end{split}
\end{equation}

For some reason these settings does not work at all with the selected optimizers, so
some research in the literature \textit{Solving Design of Pressure Vessel Engineering
Problem Using a Fruit Fly Optimization Algorithm - XIANTING KE et al} suggest the
following bounding limits:

\begin{equation}
    \begin{split}
        0.0625 \leq x_1 \leq 99 \times 0.0625\\
        0.0625 \leq x_2 \leq 99 \times 0.0625\\
        10 \leq x_3 \leq 200\\
        10 \leq x_4 \leq 200
    \end{split}
\end{equation}

But these settings produce many random, bizarre and noisy results in all selected optimizers. Then
a proud-and-lame-homemade set of variable boundings comes in handy, obtained by tweaking the original
boundings:

\begin{equation}
    \begin{split}
        0.75 \leq x_1 \leq 0.8\\
        0.35 \leq x_2 \leq 0.4\\
        39.5 \leq x_3 \leq 41.0\\
        195.0 \leq x_4 \leq 205.0\\
    \end{split}
\end{equation}

It is not intended here to point out modeling errors of any kind, nor point out package
errors made by the authors or ours, but the homemade bounding limits was necessary to
reach the literature results.

\subsection{Constraint Implementation}
\label{subsec:methodology-constraint-implementation}
The constraints that all problems are subjected were implemented as \textit{penalty functions},
that is, it adds a high value when the constraint is not satisfied, zero otherwise. It is an optional requirement
from computational work in question and is the recommended way to put constraints from \textit{MealPy}'s
Manual.

\subsection{Programming Language}
\label{subsec:methodology-programming-language}

@@

\subsection{Algorithm Selection}
\label{subsec:methodology-algorithm-selection}

The algorithm selection was made in two steps: first, it was extracted simple version of the two classes (evolutionary and swarm based) and
then it was used a simple shuffle using Python's \textit{random.shuffle} in a terminal - no script was required.

The chosen algorithms were:

\begin{itemize}
    \item \textit{\textbf{Evolution Based Algorithms: }} LevyEP (Evolutionary Programming),
    OriginalES (Evolution Strategy) and BaseGA (Genetic Algorithm)
    \item \textit{\textbf{Swarm Based Algorithms: }} OriginalBeesA (Bees Algorithm),
    OriginalFFA (Original Firefly Algorithm) and OriginalPSO (Particle Swarm Optimization)
\end{itemize}

\subsection{Optimizer Tuning}
\label{subsec:methodology-optimizer-tuning}

The hyperparameters for each optimizer were tuned using the \textit{MealPy}'s \textit{Tuner} utility.
It is a recomended procedure to tune hyperparameters for each optimizer and problem, according to
\textit{MealPy}'s Manual.
The \textit{Tuner} utility is a very simple grid search metaheuristic search tool
that test each grid configuration for a specified oprimizer runs.
Although simple, it is a very expensive procedure that took 2 days to complete. It was defined 10 runs for each
configuration, with the following set of hyperparameter:

\begin{itemize}
    \item \textit{\textbf{Evolution Based Algorithms: }}
        \subitem \textit{LevyEP:}
            \subsubitem \textit{bout\_size (float)}: percentage of child agents implement tournament selection.
        \subitem \textit{OriginalES:}
            \subsubitem \textit{lamda (float)}: Percentage of child agents evolving in the next generation.
        \subitem \textit{BaseGA:}
            \subsubitem \textit{pc (float)}: cross-over probability
            \subsubitem \textit{pm (float)}: mutation probability
    \item \textit{\textbf{Swarm Based Algorithms: }}
        \subitem \textit{OriginalBeesA:}
            \subsubitem \textit{selected\_site\_ratio (float)}
            \subsubitem \textit{elite\_site\_ratio (float)}
            \subsubitem \textit{selected\_site\_bee\_ratio (float)}
            \subsubitem \textit{elite\_site\_bee\_ratio (float)}
            \subsubitem \textit{dance\_radius (float)}
            \subsubitem \textit{dance\_reduction (float)}
        \subitem \textit{OriginalFFA:}
            \subsubitem \textit{gamma (float)}: Light Absorption Coefficient
            \subsubitem \textit{beta\_base (float)}: Attraction Coefficient Base Value
            \subsubitem \textit{alpha (float)}: Mutation Coefficient
            \subsubitem \textit{alpha\_damp (float)}: Mutation Coefficient Damp Rate
            \subsubitem \textit{delta (float)}: Mutation Step Size
            \subsubitem \textit{exponent (int)}: Exponent
        \subitem \textit{OriginalPSO:}
            \subsubitem \textit{c1 (float)}: local coefficient
            \subsubitem \textit{c2 (float)}: global coefficient
            \subsubitem \textit{w\_min (float)}: Weight min of bird
            \subsubitem \textit{w\_max (float)}: Weight max of bird

\end{itemize}

\subsection{Optimizer Parameters}
\label{subsec:methodology-optimizer-parameters}

For all optimizers and problems, were selected the following parameters:

@@

### 2023-01-19 latex/includes/header.tex
header.tex done

\subsection{Algorithm Selection}

@@

\subsection{Model Parameters}

For all models and problems, were selected the following parameters:

>>>>

\subsection{Algorithm Selection}
\label{subsec:methodology-algorithm-selection}

@@

The chosen algorithms were:

\begin{itemize}
    \item \textit{\textbf{Evolution Based Algorithms: }} LevyEP (Evolutionary Programming),
    OriginalES (Evolution Strategy) and BaseGA (Genetic Algorithm)
    \item \textit{\textbf{Swarm Based Algorithms: }} OriginalBeesA (Bees Algorithm),
    OriginalFFA (Original Firefly Algorithm) and OriginalPSO (Particle Swarm Optimization)
\end{itemize}

\subsection{Optimizer Tuning}
\label{subsec:methodology-optimizer-tuning}

The hyperparameters for each optimizer were tuned using the \textit{MealPy}'s \textit{Tuner} utility.
It is a recomended procedure to tune hyperparameters for each optimizer and problem, according to
\textit{MealPy}'s Manual.
The \textit{Tuner} utility is a very simple grid search metaheuristic search tool
that test each grid configuration for a specified oprimizer runs.
Although simple, it is a very expensive procedure that took 2 days to complete. It was defined 10 runs for each
configuration, with the following set of hyperparameter:

\begin{itemize}
    \item \textit{\textbf{Evolution Based Algorithms: }}
        \subitem \textit{LevyEP:}
            \subsubitem \textit{bout\_size (float)}: percentage of child agents implement tournament selection.
        \subitem \textit{OriginalES:}
            \subsubitem \textit{lamda (float)}: Percentage of child agents evolving in the next generation.
        \subitem \textit{BaseGA:}
            \subsubitem \textit{pc (float)}: cross-over probability
            \subsubitem \textit{pm (float)}: mutation probability
    \item \textit{\textbf{Swarm Based Algorithms: }}
        \subitem \textit{OriginalBeesA:}
            \subsubitem \textit{selected\_site\_ratio (float)}
            \subsubitem \textit{elite\_site\_ratio (float)}
            \subsubitem \textit{selected\_site\_bee\_ratio (float)}
            \subsubitem \textit{elite\_site\_bee\_ratio (float)}
            \subsubitem \textit{dance\_radius (float)}
            \subsubitem \textit{dance\_reduction (float)}
        \subitem \textit{OriginalFFA:}
            \subsubitem \textit{gamma (float)}: Light Absorption Coefficient
            \subsubitem \textit{beta\_base (float)}: Attraction Coefficient Base Value
            \subsubitem \textit{alpha (float)}: Mutation Coefficient
            \subsubitem \textit{alpha\_damp (float)}: Mutation Coefficient Damp Rate
            \subsubitem \textit{delta (float)}: Mutation Step Size
            \subsubitem \textit{exponent (int)}: Exponent
        \subitem \textit{OriginalPSO:}
            \subsubitem \textit{c1 (float)}: local coefficient
            \subsubitem \textit{c2 (float)}: global coefficient
            \subsubitem \textit{w\_min (float)}: Weight min of bird
            \subsubitem \textit{w\_max (float)}: Weight max of bird

\end{itemize}

\subsection{Optimizer Parameters}
\label{subsec:methodology-optimizer-parameters}

For all optimizers and problems, were selected the following parameters:

### 2023-01-19 latex/includes/pressure_vessel_problem.tex
header.tex done

\begin{table}
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &        Mean F &      Median F &          Max F &      StdDev F \\
\midrule
       EP & 0.94974179 &    0.99959274 &    0.95597819 &     1.41652156 &    0.10256309 \\
       ES & 0.94985828 &    3.07906532 &    1.90278615 &    11.41620650 &    2.47339648 \\
       GA & 4.61928311 &  383.23040981 &  262.27555979 &  1747.41758530 &  362.98174772 \\
    BeesA & 1.05944833 & 5212.47847291 & 2094.42824655 & 30412.73885965 & 7297.64717603 \\
      FFA & 0.95407161 &    1.05749808 &    1.05432826 &     1.16958618 &    0.04485239 \\
      PSO & 0.94936617 &    1.00652465 &    0.94974267 &     3.01988559 &    0.25222125 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{table}
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &        Mean F &      Median F &         Max F &   StdDev F \\
\midrule
       EP & 5534.57345496 & 5534.63426098 & 5534.61416291 & 5534.92993607 & 0.05804548 \\
       ES & 5534.57917966 & 5535.28165367 & 5535.06149254 & 5538.03003895 & 0.65098467 \\
       GA & 5534.83662424 & 5538.36788889 & 5536.92932748 & 5550.91639342 & 3.57079951 \\
    BeesA & 5534.57516657 & 5537.97495231 & 5537.04824684 & 5548.55587234 & 3.28239740 \\
      FFA & 5534.57401617 & 5534.58849395 & 5534.58435493 & 5534.63296149 & 0.01341777 \\
      PSO & 5534.57345480 & 5534.57381593 & 5534.57355270 & 5534.57625290 & 0.00053848 \\
\bottomrule
\end{tabular}
\end{table}

### 2023-01-19 latex/includes/spring_problem.tex
header.tex done

\begin{table}
\centering
\caption{Statistical Information about function values for Spring Tension Design}
\label{function_values:spring_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &     Mean F &   Median F &      Max F &   StdDev F \\
\midrule
       EP & 0.00607563 & 0.00607590 & 0.00607570 & 0.00607834 & 0.00000044 \\
       ES & 0.00607564 & 0.00609326 & 0.00608915 & 0.00616308 & 0.00001810 \\
       GA & 0.00610989 & 0.00670843 & 0.00663331 & 0.00805670 & 0.00045944 \\
    BeesA & 0.00607781 & 0.00632310 & 0.00625074 & 0.00724816 & 0.00025469 \\
      FFA & 0.00607563 & 0.00607593 & 0.00607576 & 0.00607847 & 0.00000047 \\
      PSO & 0.00607563 & 0.00607563 & 0.00607563 & 0.00607563 & 0.00000000 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{table}
\centering
\caption{Statistical Information about function values for Spring Tension Design}
\label{function_values:spring_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &     Mean F &   Median F &      Max F &   StdDev F \\
\midrule
       EP & 0.01250534 & 0.01255960 & 0.01253156 & 0.01453933 & 0.00020190 \\
       ES & 0.01252956 & 0.01281857 & 0.01274283 & 0.01346214 & 0.00024181 \\
       GA & 0.01252305 & 0.01628575 & 0.01616768 & 0.02561808 & 0.00273276 \\
    BeesA & 0.01250200 & 0.01268010 & 0.01261691 & 0.01406679 & 0.00021629 \\
      FFA & 0.01249803 & 0.01254780 & 0.01253499 & 0.01268116 & 0.00003152 \\
      PSO & 0.01252626 & 0.01390866 & 0.01318774 & 0.03036537 & 0.00342240 \\
\bottomrule
\end{tabular}
\end{table}
